A Case Study in Preserving a\\High Energy Physics Application

### 2014-03-14 paper.tex
Edited the evolution section.

\title{A Case Study in Preserving a High Energy Physics Application}
\author{
Haiyan Meng, Matthias Wolf, Peter Ivie, Anna Woodard, Michael Hildreth, and Douglas Thain\\
\affaddr{Department of Physics and Department of Computer Science and Engineering}\\
\affaddr{University of Notre Dame}\\
\affaddr{\{hmeng|mwolf3|pivie|awoodard|mhildreth|dthain\}@nd.edu}
}
\date{15 January 2014}
\maketitle

@@

\section{Preservation Strategies}

Figure~\ref{fig:version-evolution} illustrates the evolution history of
different preservation strategies. First, an email depicting how to repeat the
experiment is used. Then, all the necessary steps to reproduce the experiment
mentioned in the mail are collected into one shell script. In order to decouple
data sources and data references, one formulated script is introduced
to express all the dependencies at the beginning of the script in the form of
environment variables. Finally, this decoupling idea is further extended into
one map file which keeps the relationship between the data reference in the
script and the actual storage location. One independent package containing all
the data and software dependencies is generated to make the reproduction
process more easier.

\subsection{Solution 1: Email}

In order to repeat Matthais's example, we consulted the original
author by email about the necessary work for the experiment. In response,
the original author introduced the general workflow of his tau roast program through one long
email including notes, linux shell commands, web linksi, as illustrated by the Version 1 of
Table~\ref{table:scripts}.

However, this solution to repeat one experiment has three potential drawbacks.
Firstly, the experience of repeating one tau roast program through emails is
chaotic. You need to constantly jump around multiple web links. There are overlap
between the content of the email and the content of web links, which needs the
new user to merge them. Multiple communication through emails is necessary to
ensure the successful reproduction of the whole analysis. For example, the original
email refers to one environment variable called CMSSW\_base without clear
declaration, the new user needs to send one email to the original writer to
obtain its accurate value. Secondly, the necessary procedures to repeat the
experiment, including software acquisition from different sources, is complex
for the new user. In the example, the sources of software includes
CMSSW, Git, HTTP. The access of CMSSW requires the new user to be an authorized
user of CMSSW. What's more, some parts of the workflow are unrepeatable. The
third step of this experiment requires the new user to own the access authority
of the Grid and involves the analysis of AOD data with the size of about 400TB.

Implication: Directly repeating the experiment using the workflow description and results provided by the original author is complex and difficult. Some extra work must be done to make the reproduction process easier.

\subsection{Solution 2: Script} One possible solution for the access authority
problem of Grid data is to grant the new user the authority to access gird data
directly. Another possible solution is let the new user directly operate on the
machine the original author used. As for the complexity of jumping between
multiple web links, people may suggest that the original author
should generate one script including all the contents from different web links.

To test out these possible solutions, we integrate the content of all the notes, commands and web links involved in the emails into one neat, complete shell script, which begins with the definition of environment variables, software acquisition from CMSSW, Git and other web resources, and software installation, ends with the execution of the actual analysis program. Version 2 of Table~\ref{table:scripts} illstrates the merged script. As for the data acquisition from the Grid, we directly use the local copy in HDFS to avoid requiring the new user to obtain a Grid certification.

%*****hmeng-doult: the size of hadoop need to reconsidered. how to correctly illustrate the data size of this example?

Because the script includes all the necessary procedures for the reproduction
of one analysis program, the readability and friendliness of this solution is
higher than that of the email format.

\subsection{Solution 3: Formulated Script}
%%

The script can reduce the complexity of repeating one experiment through the integration of all the necessary procedures. However, the data and software dependencies are still randomly distributed across the whole script, which requires one complete scan of the script to obtain the  dependency list.

To decouple data sources and data references and make the data dependencies clear, Version 2 of Table~\ref{table:scripts} is formulated into one new version, as shown in
the Version 3 of Table~\ref{table:scripts}. Each new environment variable at the beginning part of
the script is corresponding to one dependency. All the following access to data
or software dependency will refer to its corresponding environment variable.
All the environment variables of dependencies form one map, which maintains the
target address of each dependency.

This script style may function as one new guideline to the original author of
one experiment, which expresses the dependencies more clearly and expedites the
preparation process to repeat one experiment. The introduction of the map file
also reduces the workload of changing one dependency. For example, if we want
to utilize one new git package to analyze the same dataset, only the modification of
environment variable corresponding to the original git package is necessary.
Without the map mechanism, the whole script needs to be scanned to figure out and
replace all the references of the original git package.

So far, the data access authority is ignored and
the reproduction of the experiment is on the machine the original author used. 
As for Grid data, using the data
copy stored in the machine where the original author executed the experiment requires the
new user to have the authority to access the machine, which complicates the
management of the original machine, even is impossible if the original machine
executes rigid user access control. Granting everyone who wants to repeat the
analysis the access authority for Grid is unacceptable to system administrator.

Implication: All the data and software involved in the experiment should be
provided to the new user in the format of one self-contained package so that
the new user can avoid 
access authority problems.
In addition, the requirements of the underlying
OS and hardware should also be figured out and provided to the new user.

>>>>

\title{A Case Study in Preserving a\\High Energy Physics Application}
\author{
Haiyan Meng, Matthias Wolf, Peter Ivie, Anna Woodard, Michael Hildreth, and Douglas Thain\\
\affaddr{Department of Physics and Department of Computer Science and Engineering}\\
\affaddr{University of Notre Dame}\\
\affaddr{\{hmeng|mwolf3|pivie|awoodard|mhildreth|dthain\}@nd.edu}
}
\date{15 January 2014}
\maketitle

@@

\section{Evolving the Artifact}

It is clear that the artifact, as provided, is not in a suitable form
for preservation.  While it might be technically possible to automatically
capture the entire virtual machine and all of the connected filesystems,
doing so would be wasteful and prohibitively expensive.  Further, if multiple
similar applications are preserved, we would miss the opportunity to identify
common dependencies and store them once for multiple artifacts.

Figure~\ref{fig:version-evolution} shows how we have evolved this artifact
through several stages which make it more suitable for preservation.
In each step of evolution, we make the dependencies of the artifact
more explicit and available for analysis and automated processing.
As noted in the previous section, the original author provided us with
prose instructions by email which we translated into an
executable script.  The executable script has embedded in it
a number of external identifiers such as URLs pointing to repositories
and paths to networked filesystems.  As a general programming practice,
embedding such constants into the middle of a program is unwise, and so
we extract all of those identifiers and place them at the beginning
of the script in the form of variable substitutions.  This makes
it readily apparent to the reader what the dependencies are.

Form 4 goes a step further and takes all of the dependencies \emph{out}
of the script itself and places them in a \emph{dependency map}.
The dependency map lists all of the external dependencies of the application, indicating
the type, how they are accessed, and where they are currently located.
The resulting script then simply refers to abstract file locations such
as \verb!$CMSSW!, while the map file indicates where they are located.
If properly constructed, the script should not refer to any external
resource unless it is indicated in the dependency map.

By extracting the dependencies into the external dependency map,
we introduce great freedom for the curator to move, transform, and otherwise
manipulate the dependencies of the artifact without damaging the artifact itself.
For example, it is straightforward for an automated tool to examine
all of the dependencies in the map, download those that are missing,
and then modifiy the map to point to the local copies of the dependencies.

\section{Fine Grained Dependency Management}
